Attract and kill: Spinosad containing spheres to control onion maggot (\textit{Delia antiqua})

### 2020-01-21 manuscript/pest-management-science/main.tex
Updates from Overleaf

Using spinosad containing spheres as an attract and kill solution for control of \textit{D. antiqua} populations may be an effective integrated pest management tool.  Over the course of the three field seasons, with three generations per year (Figure \ref{fig:figure1}), the attract and kill solution was able to consistently inflict mortality on adult \textit{D. antiqua} flies, even after prolonged exposure to field conditions (Figure \ref{fig:figure2}).  While prolonged field exposure did tend to reduce efficacy (Figure \ref{fig:figure2}C) and change the dynamics of mortality (Figure \ref{fig:figure3}), any effects were marginal and may be due to melting of the paraffin solution containing the spinosad (Figure \ref{fig:figure2}D).

@@

Using \textit{D. antiqua} population averages from field experiments and mortality achieved by field-exposed spheres over time assessed in the laboratory, numbers of adult \textit{D. antiqua}  killed as a result of the attract and kill strategy using spinosad can be estimated. During peak flights, approximately 272 \textit{D. antiqua} adults (both males and females) were caught on average per week.  Over the course of the field season, approximately 64 adults were caught by sticky cards on average per week.  Assuming spinosad containing spheres kill approximately 54\% of the adults they contact, over the course of a 16 week season, this attract and kill solution could have killed approximately 553 \textit{D. antiqua} adults (16 weeks x 64 adults on average per week * 0.54 mortality) per spinosad containing sphere.

Comparisons of attract and kill strategies with insecticide and no insecticide controls suggest that spinosad containing spheres are reducing some damage by \textit{D. antiqua} larvae.  Spinosad spheres in conjunction with the attractive Delia Lure resulted in levels of onion damage almost as good as insecticide controls, but with variation that led to overlap with no insecticide negative controls.  In caged field trials, spinosad containing spheres reduced damage resulting from \textit{D. antiqua} feeding to below that of no insecticide controls.  Interestingly, pairing Delia Lure with sticky cards produced higher levels of damage.  This is likely because the kill part of the attract and kill sticky card solution was ineffective; Delia Lure likely attracted adult fly populations to plots but sticky cards did not cause sufficient mortality.  Augmenting spinosad spheres with ammonium carbonate and casein hydrolysate did not appreciably change control with this attract and kill method.

>>>>

Using spinosad containing spheres as an attract and kill solution for control of \textit{D. antiqua} populations may be an effective integrated pest management tool.  Over the course of the three field seasons, with three generations per year (Figure \ref{fig:figure1}), the attract and kill solution was able to consistently inflict mortality on adult \textit{D. antiqua} flies, even after prolonged exposure to field conditions (Figure \ref{fig:figure2}).  While prolonged field exposure did tend to reduce efficacy (Figure \ref{fig:figure2}C) and change the dynamics of mortality (Figure \ref{fig:figure3}), any effects were marginal and may be due to melting (personal observation) of the paraffin solution containing the spinosad (Figure \ref{fig:figure2}D).

@@

Numbers of adult \textit{D. antiqua}  killed as a result of the attract and kill strategy using spinosad can be estimated using \textit{D. antiqua} population averages from field experiments and mortality achieved by field-exposed spheres over time assessed in the laboratory.  During peak flights, approximately 272 \textit{D. antiqua} adults (both males and females) were caught by sticky cards on average per week.  Over the course of the field season, approximately 64 adults were caught by sticky cards on average per week.  Assuming spinosad containing spheres kill approximately 54\% of the adults they contact, over the course of a 16 week season, this attract and kill solution could have killed approximately 553 \textit{D. antiqua} adults (16 weeks x 64 adults on average per week * 0.54 mortality) per spinosad containing sphere.

Comparisons of attract and kill strategies with insecticide and no insecticide controls suggest that spinosad containing spheres are reducing some damage by \textit{D. antiqua} larvae.  Spinosad spheres in conjunction with the attractive Delia Lure resulted in levels of onion damage almost as good as insecticide controls, but with variation that led to overlap with no insecticide negative controls.  In caged field trials, spinosad containing spheres reduced damage resulting from \textit{D. antiqua} feeding to below that of no insecticide controls.  Interestingly, pairing Delia Lure with sticky cards produced higher levels of damage.  This is likely because the kill part of the attract and kill sticky card solution was ineffective; Delia Lure likely attracted adult fly populations to plots but sticky cards did not cause sufficient mortality.  Augmenting spinosad spheres with ammonium carbonate and casein hydrolysate, added to increase longevity of the paraffin,  did not appreciably change control with this attract and kill method.
